---
title: "Machine Learning PGA MASTERS"
author: "Sam Milnes"
date: "3/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

pga <- read.csv("C:\\Users\\milness\\Desktop\\Machine Learning\\Final Project\\2022Masters_pred_data.csv")
pgaPast5 <- read.csv("C:\\Users\\milness\\Desktop\\Machine Learning\\Final Project\\2022Masters_past5.csv")

# pga <- pga[,-1]
# 
# 
# pga <- subset(pga, !is.na(pga$masters_finish))


#sub2021 <- pga[which(pga$year == 2021), ]
#View(sub2021)
```

DEFINING COLUMN NAMES:


Top10 -> How many top 10 finishes did the player have for that given year

Wins -> How many tour wins did the player have for that given year

Score Average -> The average score the player had for given year

Rounds -> How many rounds the player played that given year

Bounce Back -> Percent of holes that scored over par immediately followed by a hole that scored less than par

Driving Accuracy -> The percentage of time a tee shot comes to rest in the fairway (regardless of club)

Driving Distance -> The average number of yards per measured drive

Gir -> (Greens in regulation) The percent of time a player was able to hit the green in regulation (greens hit in regulation/holes played)

Hole_Proximity -> The average distance the ball comes to rest from the hole (in feet) after the player's approach shot.

x_of_attempts -> 

putts_round ->  The average number of total putts per round

scramble -> Percent that shows a players ability to make par (or better) even if a Green in Regulation (GIR) is missed

sg_putt -> The number of putts a player takes from a specific distance is measured against a statistical baseline to determine the player's strokes gained or lost on a hole.

sg_t2g -> The per round average of the number of Strokes the player was better or worse than the field average on the same course & event minus the Players Strokes Gained putting value.

sg_total -> The per round average of the number of Strokes the player was better or worse than the field average on the same course & event.

sg_approach -> The number of Approach the Green strokes a player takes from specific locations and distances are measured against a statistical baseline to determine the player's strokes gained or lost on a hole

sg_tee -> The number of strokes a player takes from a specific distance off the tee on Par 4 & par 5's is measured against a statistical baseline to determine the player's strokes gained or lost off the tee on a hole

sg_ar -> The number of Around the Green strokes a player takes from specific locations and distances are measured against a statistical baseline to determine the player's strokes gained or lost on a hole

par4_sa -> 

par3_sa ->

par5_sa -> 

covered_tee -> 

left_rough ->

right_rough -> 

fair_bunk -> 

gir_fair -> 

gir_not_fair -> 

prox_fair ->

prox_rough -> 

go_for_green ->

sand_save -> 

prox_arg ->

putt_25 ->

putt15_25 ->

putt10_15 ->

putt_10 ->

finround_sa ->

masters_finish ->

total_score -> 




List of players playing in 2022 Masters
```{r}

masters2022_list = list("Abraham Ancer",
                     "Daniel Berger",
                     "Christiaan Bezuidenhout",
                     "Sam Burns",
                     "Patrick Cantlay",
                     "Paul Casey",
                     "Cameron Champ",
                     "Stewart Cink",
                     "Corey Conners",
                     "Fred Couples",
                     "Cameron Davis",
                     "Bryson DeChambeau",
                     "Harry English",
                     "Tony Finau",
                     "Matthew Fitzpatrick",
                     "Tommy Fleetwood",
                     "Sergio Garcia",
                     "Lucas Glover",
                     "Talor Gooch",
                     "Austin Greaser",
                     "Steward Hagestad",
                     "Brian Harman",
                     "Padraig Harrington",
                     "Tyrrell Hatton",
                     "Lucas Herbert",
                     "Garrick Higgo",
                     "Harry Higgs",
                     "Tom Hoge",
                     "Max Homa",
                     "Billy Horschel",
                     "Viktor Hovland",
                     "Mackenzie Hughes",
                     "Sungjae Im",
                     "Aaron Jarvis",
                     "Dustin Johnson",
                     "Zach Johnson",
                     "Takumi Kanaya",
                     "Si Woo Kim",
                     "Kevin Kisner",
                     "Brooks Koepka",
                     "Jason Kokrak",
                     "Bernhard Langer",
                     "Kyoung-Hoon Lee",
                     "Min Woo Lee",
                     "Marc Leishman",
                     "Luke List",
                     "Shane Lowry",
                     "Sandy Lyle",
                     "Robert MacIntyre",
                     "Hideki Matsuyama",
                     "Rory McIlroy",
                     "Guido Migliozzi",
                     "Larry Mize",
                     "Francesco Molinari",
                     "Collin Morikawa",
                     "Kevin Na",
                     "Keita Nakajima",
                     "Joaquin Niemann",
                     "Jose Mario Olazabal", #Check for spelling
                     "Louis Oosthuizen",
                     "Ryan Palmer",
                     "James Piot",
                     "Jon Rahm",
                     "Patrick Reed",
                     "Justin Rose",
                     "Xander Schauffele",
                     "Scottie Scheffler",
                     "Charl Schwartzel",
                     "Adam Scott",
                     "Laird Shepherd",
                     "Webb Simpson",
                     "Vijay Singh",
                     "Camerson Smith",
                     "Jordon Spieth",
                     "Sepp Straka",
                     "Hudson Swafford",
                     "Justin Thomas",
                     "Erik van Rooyen",
                     "Bubba Watson",
                     "Mike Weir",
                     "Lee Westwood",
                     "Danny Willet",
                     "Matthew Wolff",
                     "Gary Woodland",
                     "Tiger Woods",
                     "Will Zalatoris")
```

Filter data to only players playing in Masters this year.
```{r}
library(dplyr)    

current_masters <- filter(pga, player_name %in% masters2022_list)

current_masters <- current_masters[, -1]
View(current_masters)

```

